---
title: "Histogram_circle_count"
author: "Henry"
date: '2022-09-07'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following code generates histograms/density plots that display the size distribution of a dataset

## Gather data from multiple tab files containing the name and sizes of all circles (plasmids + phages) into one dataframe
```{r}
getwd()

# get list of files
files <- list.files(path = "histogram_sample_circles_count", pattern = ".txt", full.names = TRUE, recursive = FALSE)

# read table on list of files
df <- lapply(files, function(x) {
  a <- read.table(x, col.names = c("contig", "bases"), colClasses = c("character", "numeric"))
})

# rename df's to corresponding file names
names(df) <- list.files(path = "histogram_sample_circles_count", pattern = ".txt", full.names = TRUE, recursive = FALSE)

# at this point we have a list of dataframes, where each dataframe is represents a sample. When plotting, you select which dataframe to use by index (i.e. df[1] will be sample 1 wheras df[2] will be sample 2)
list2env(df, envir = .GlobalEnv)
```

## Generates other tables for individual subsets (plasmid + phage, plasmid only, phage only, recurring, singletons)
```{r}
# All circles
all_circles_sizes <- read.table("histogram_sample_circles_count/recurring_vs_singletons/all_contigs_plasmid_sizes.txt")

# All phages
phages <- read.table("histogram_sample_circles_count/recurring_vs_singletons/all_phages_sizes.txt")

# All recurring (plasmids only)
recurring_plasmid <- read.table("histogram_sample_circles_count/recurring_vs_singletons/recurring_plasmid_sizes.txt")

# All recurring (phages)
recurring_phage <- read.table("histogram_sample_circles_count/recurring_vs_singletons/recurring_phages_sizes.txt")

# Combined recurring (plasmid + phage)
all_recurring <- rbind(recurring_plasmid, recurring_phage)

# Singletons (includes plasmid and phages)
singles <- all_circles_sizes[!(all_circles_sizes$V1 %in% all_recurring$V1), ]

# Single phages
single_phage <- phages[!(phages$V1 %in% recurring_phage$V1), ]

# Single plasmids
single_plasmid <- singles[!(singles$V1 %in% single_phage), ]
```

## Create ggplot2 histogram

4 panel for singleton and recurring plasmids and phages
```{r}
library(ggplot2)
library(gridExtra)
library(ggpubr)

# multiply y axis values
mult_format <- function() {
  function(x) format(100000 * x, digits = 1)
}

# display x axis without scientific notation
options(scipen = 999)

hist.p1 <- ggplot() +
  geom_density(
    data = single_plasmid, aes(x = V2),
    alpha = 0.5,
    # blue
    fill = "#39B185",
    color = "#39B185",
    size = 1
  ) +
  # ggtitle("Singleton Plasmids") +
  scale_x_continuous(
    limits = c(0, 100000),
    breaks = c(seq(0, 100000, by = 10000)),
    labels = c(seq(0, 90, by = 10), "100"),
    expand = c(0, 0),
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    labels = mult_format(),
    limits = c(0, 0.000025),
  ) +
  labs(x = "Size (kbs)", y = bquote("Density "(1 ~ x ~ 10^-5))) +
  theme(
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none",
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 20, face = "bold"),
    plot.margin = margin(10, 30, 10, 10)
  )

hist.p2 <- ggplot() +
  geom_density(
    data = recurring_plasmid, aes(x = V2),
    alpha = 0.5,
    # blue
    fill = "#39B185",
    color = "#39B185",
    size = 1
  ) +
  # ggtitle("Recurring Plasmids") +
  scale_x_continuous(
    limits = c(0, 100000),
    breaks = c(seq(0, 100000, by = 10000)),
    labels = c(seq(0, 90, by = 10), "100"),
    expand = c(0, 0),
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    labels = mult_format()
  ) +
  labs(x = "Size (kbs)", y = bquote("Density "(1 ~ x ~ 10^-5))) +
  theme(
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none",
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 20, face = "bold"),
    plot.margin = margin(10, 30, 10, 10)
  )
hist.p3 <- ggplot() +
  geom_density(
    data = single_phage, aes(x = V2),
    alpha = 0.5,
    # blue
    fill = "#FFC61E",
    color = "#FFC61E",
    size = 1
  ) +
  # ggtitle("Singleton Phage") +
  scale_x_continuous(
    limits = c(0, 100000),
    breaks = c(seq(0, 100000, by = 10000)),
    labels = c(seq(0, 90, by = 10), "100"),
    expand = c(0, 0),
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    labels = mult_format(),
    limits = c(0, 0.00008),
  ) +
  labs(x = "Size (kbs)", y = bquote("Density "(1 ~ x ~ 10^-5))) +
  theme(
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none",
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 20, face = "bold"),
    plot.margin = margin(10, 30, 10, 10)
  )

hist.p4 <- ggplot() +
  geom_density(
    data = recurring_phage, aes(x = V2),
    alpha = 0.5,
    # blue
    fill = "#FFC61E",
    color = "#FFC61E",
    size = 1
  ) +
  # ggtitle("Recurring Phage") +
  scale_x_continuous(
    limits = c(0, 100000),
    breaks = c(seq(0, 100000, by = 10000)),
    labels = c(seq(0, 90, by = 10), "100"),
    expand = c(0, 0),
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    labels = mult_format(),
    limits = c(0, 0.00008)
  ) +
  labs(x = "Size (kbs)", y = bquote("Density "(1 ~ x ~ 10^-5))) +
  theme(
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    legend.position = "none",
    axis.title = element_text(size = 25, face = "bold"),
    axis.text = element_text(size = 20, face = "bold"),
    plot.margin = margin(10, 30, 10, 10)
  )

ggarrange(hist.p1, hist.p2, hist.p3, hist.p4,
  ncol = 2,
  nrow = 2,
  labels = c("A", "B", "C", "D"),
  font.label = list(size = 25)
)

ggsave
```

All sample densities overlaid
```{r}
library(ggplot2)

# display x axis without scientific notation
options(scipen = 999)

all.dens <- ggplot() +
  geom_density(
    data = subset(df[[1]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF",
  ) +
  geom_density(
    data = subset(df[[2]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF"
  ) +
  geom_density(
    data = subset(df[[3]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF"
  ) +
  geom_density(
    data = subset(df[[4]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF"
  ) +
  geom_density(
    data = subset(df[[5]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF"
  ) +
  geom_density(
    data = subset(df[[6]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF"
  ) +
  geom_density(
    data = subset(df[[7]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF"
  ) +
  geom_density(
    data = subset(df[[8]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF"
  ) +
  geom_density(
    data = subset(df[[9]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF"
  ) +
  geom_density(
    data = subset(df[[10]], bases < 100000), aes(x = bases),
    fill = "#0074FF40",
    color = "#0074FF"
  ) +
  scale_x_continuous(
    limits = c(0, 100000),
    breaks = c(seq(0, 100000, by = 10000)),
    labels = c(seq(0, 90, by = 10), "100"),
    expand = c(0, 0)
  ) +
  scale_y_continuous(
    expand = c(0, 0),
    breaks = c(seq(0, 0.00004, by = 0.00001)),
    limits = c(0, 0.00004),
    labels = c(seq(0, 3, by = 1), "4")
  ) +
  labs(x = "Size (kbs)", y = expression("Density "(1 ~ x ~ 10^-5))) +
  theme(
    panel.background = element_blank(),
    axis.line = element_line(colour = "black"),
    plot.margin = margin(10, 30, 10, 10),
    text = element_text(size = 25, face = "bold"),
    axis.title = element_text(size = 25)
  )

# include recurring, singleton plas/phage in a 5 figure

ggarrange(all.dens,
  ggarrange(hist.p1, hist.p2,
    ncol = 2,
    labels = c("B", "C"),
    font.label = c(size = 35),
    vjust = 1.1
  ),
  ggarrange(hist.p3, hist.p4,
    ncol = 2,
    labels = c("D", "E"),
    font.label = c(size = 35),
    vjust = 1.1
  ),
  nrow = 3,
  heights = c(2, 1, 1),
  labels = c("A"),
  font.label = c(size = 35)
)
```
