---
title: "stripchart"
author: "Henry"
date: '2022-12-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Initialize data for number of annotations containing KO/GO pathways
```{r}
# get list of files in current working directory
files <- list.files(path = "./with_terms/", pattern = ".txt", full.names = TRUE, recursive = FALSE)

# read table on list of files to create a list of dataframes
df <- lapply(files, function(x) {
  a <- read.table(x, col.names = c("sample", "num_cds", "bakta_pred", "foldseek_pred_GO", "foldseek_pred_KO"), colClasses = c("character", "numeric", "numeric", "numeric", "numeric"))
})

# rename df's to its sample name (without the extension and _num_predictions suffix)
names(df) <- gsub("_num_path_annos", "", tools::file_path_sans_ext(list.files(path = "./with_terms", pattern = ".txt", recursive = FALSE)))

list2env(df, envir = .GlobalEnv)
```

Create new dataframe with all data in one dataframe for plotting
```{r}
# DF FOR ALL SAMPLES
all <- data.frame(Sample = character(), Percent = numeric(), Method = character())

# iterate over list of dataframes (samples)
for (i in 1:length(df)) {
  y <- data.frame(df[i])
  # iterate over each row, append to total dataframe
  for (x in 1:nrow(y)) {
    all[nrow(all) + 1, ] <- c(names(df[i]), (y[x, 3]) / (y[x, 2]), "Bakta")
    all[nrow(all) + 1, ] <- c(names(df[i]), ((y[x, 4]) + y[x, 5]) / (y[x, 2]), "Colabfold/Foldseek")
  }
}

all["Percent"][all["Percent"] == 1] <- 0

#-----------------------------------------

# DF FOR SINGLE SAMPLE (T2_may3-21)
s <- data.frame(Sample = character(), Percent = numeric(), Method = character())

# build dataframe - calculate percent annotations
for (i in 1:nrow(`GAC6`)) {
  s[nrow(s) + 1, ] <- c("GAC6", (`GAC6`[i, 3]) / (`GAC6`[i, 2]), "Bakta")
  s[nrow(s) + 1, ] <- c("GAC6", (GAC6[i, 4] + GAC6[i,5]) / GAC6[i, 2], "Colabfold/Foldseek")
}
```


#Create single plot for percent of annotations containing KEGG/GO  (representing all samples)
```{r}
library(ggplot2)

path.anno <- ggplot(all, aes(x = Sample, y = as.numeric(Percent) * 100, color = Method)) +
  geom_jitter(position = position_jitterdodge(
    jitter.width = 0.3,
    dodge.width = 0.6
    ),
    size = 4,
    alpha = 0.5
  ) +
  #scale_y_continuous(expand = c(0,0))+
  coord_cartesian(ylim=c(0,40)
                  )+
  labs( # title="CDS' containing Pathway Annotations (GO/KEGG) by Method",
    x = "Sample", y = "Annotations with KEGG/GO Terms (%)"
  ) +
  stat_summary(
    aes(color = Method),
    linewidth = 1,
    fun.data = "mean_sdl",
    fun.args = list(mult = 1),
    shape = 18,
    size = 1.3,
    na.rm = TRUE
  ) +
  scale_color_manual(values = c("#FF507A", "#0082ff")) +
  theme(
    panel.background = element_blank(),
    axis.line = element_line(colour = "#3A3A3A"),
    axis.title = element_text(size = 13, face = "bold"),
    axis.text = element_text(size = 15),
  )
```

#Create single plot for percent of annotations containing KEGG and/or GO  (representing T2_may3)
```{r}
library(ggplot2)

# panel 2
strip.p2 <- ggplot(s, aes(x = Sample, y = as.numeric(Percent) * 100, color = Method)) +
  geom_jitter(
    position = position_jitterdodge(
      jitter.width = 0.3,
      dodge.width = 0.6
    ),
    size = 4,
    alpha = 0.5
  ) +
  scale_y_continuous(
    limits = c(0, 25),
  ) +
  labs( # title="CDS' containing Pathway Annotations (GO/KEGG) by Method",
    x = "Sample", y = "Annotations with KEGG/GO Terms (%)"
  ) +
  stat_summary(aes(color = Method),
    linewidth = 1,
    fun.data = "mean_sdl",
    fun.args = list(mult = 1),
    shape = 15,
    size = 1.3,
    na.rm = TRUE
  ) +
  scale_color_manual(values = c("#FF507A", "#0082ff")) +
  theme(
    panel.background = element_blank(),
    axis.line = element_line(colour = "#3A3A3A"),
    axis.title = element_text(size = 10, face = "bold"),
    axis.text = element_text(size = 10)
  )
```


Initialize data for number of hypotheticals 
```{r}
# get list of files in current working directory
files2 <- list.files(path = "./with_identifiable_protein/", pattern = ".txt", full.names = TRUE, recursive = FALSE)

# read table on list of files to create a list of dataframes
df2 <- lapply(files2, function(x) {
  a <- read.table(x, sep = "\t", col.names = c("contig", "num_cds", "num_hypoth", "foldseek_pred"), colClasses = c("character", "numeric", "numeric", "numeric"))
})

# rename df's to its sample name (without the extension and _num_predictions suffix)
names(df2) <- gsub("_num_hypoths", "", tools::file_path_sans_ext(list.files(path = "./with_identifiable_protein", pattern = ".txt", recursive = FALSE)))

list2env(df2, envir = .GlobalEnv)
```

Create new dataframe with all data in one dataframe for plotting
```{r}
# FOR ALL SAMPLES
all2 <- data.frame(Sample = character(), Percent = numeric(), Method = character())

# iterate over list of dataframes (samples)
for (i in 1:length(df2)) {
  y <- data.frame(df2[i])
  # iterate over each row, append to total dataframe
  for (x in 1:nrow(y)) {
    all2[nrow(all2) + 1, ] <- c(names(df2[i]), (y[x, 2] - y[x, 3]) / (y[x, 2]), "Bakta")
    all2[nrow(all2) + 1, ] <- c(names(df2[i]), ((y[x, 2]) - (y[x, 3] - y[x, 4])) / (y[x, 2]), "Colabfold/Foldseek")
  }
}

# FOR SINGLE SAMPLE
s2 <- data.frame(Sample = character(), Percent = numeric(), Method = character())
y <- `GAC6`

# calculate percent annotations
for (x in 1:nrow(`GAC6`)) {
  s2[nrow(s2) + 1, ] <- c("GAC6", (y[x, 2] - y[x, 3]) / (y[x, 2]), "Bakta")
  s2[nrow(s2) + 1, ] <- c("GAC6", ((y[x, 2]) - (y[x, 3] - y[x, 4])) / (y[x, 2]), "Colabfold/Foldseek")
}
```

#Create single plot for percent of identifiable proteins (non-hypotheticals) (representing all samples)
```{r}
library(ggplot2)

iden <- ggplot(all2, aes(x = Sample, y = as.numeric(Percent) * 100, color = Method)) +
  geom_jitter(
    position = position_jitterdodge(
      jitter.width = 0.3,
      dodge.width = 0.6
    ),
    size = 4,
    alpha = 0.5
  ) +
  #scale_y_continuous(expand = c(0,0)) +
  coord_cartesian(ylim=c(0,100)
                  )+
  labs( # title = "Percent of Identifiable Proteins by Method In Phages",
    x = "Sample", y = "Identifiable Proteins (%)"
  ) +
  stat_summary(aes(group = Method),
    linewidth = 1,
    fun.data = "mean_sdl",
    fun.args = list(mult = 1),
    shape = 18,
    size = 1.3,
    na.rm = FALSE
  ) +
  scale_color_manual(values = c("#FF507A", "#0082ff")) +
  theme(
    panel.background = element_blank(),
    axis.line = element_line(colour = "#3A3A3A"),
    axis.title = element_text(size = 13, face = "bold"),
    axis.text = element_text(size = 15)
  )
```

#Create single plot for percent of identifiable proteins (non-hypothetical) (representing T2_may3)
```{r}
library(ggplot2)

strip.p1 <- ggplot(s2, aes(x = Sample, y = as.numeric(Percent) * 100, color = Method)) +
  geom_jitter(
    position = position_jitterdodge(
      jitter.width = 0.3,
      dodge.width = 0.6
    ),
    size = 4,
    alpha = 0.5
  ) +
  scale_y_continuous(
    limits = c(0, 100),
  ) +
  labs( # title = "Percent of Identifiable Proteins by Method In Phages",
    x = "Sample", y = "Identifiable Proteins (%)"
  ) +
  stat_summary(aes(color = Method),
    linewidth = 1,
    fun.data = "mean_sdl",
    fun.args = list(mult = 1),
    shape = 18,
    size = 1.3,
    na.rm = TRUE
  ) +
  scale_color_manual(values = c("#FF507A", "#0082ff")) +
  theme(
    panel.background = element_blank(),
    axis.line = element_line(colour = "#3A3A3A"),
    axis.title = element_text(size = 10, face = "bold"),
    axis.text = element_text(size = 10)
  )
```

Side by side of plots representing all samples
```{r}
library(gridExtra)
library(ggpubr)

ggarrange(iden, path.anno,
                  ncol = 1, 
                  nrow = 2,
                  labels = c("A","B"),
                  common.legend = TRUE,
                  legend = "right"
                  )
```